---
title: "In-class Exercise 10"
description: |
  This post presents the step by step instructions that were followed to achieve the corresponding visualizations in Session 10 of our Visual Analytics Module.
author:
  - name: Mayurapriyann Arulmozhi Baskaran
    url: https://www.linkedin.com/in/mayurapriyann/
date: 07-11-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
    smooth_scroll: true
    theme: united
    highlight: tango
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, 
                      echo = TRUE, 
                      eval = TRUE,
                      message = FALSE, 
                      error = FALSE, 
                      warning=FALSE)
```

```{r}
packages = c('tidytext','widyr','wordcloud','DT','ggwordcloud',
             'textplot','lubridate','dplyr','hms','tidyverse','tidygraph',
             'ggraph','igraph')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```
# Creating a folder list

```{r}
news20 <- "C:/Users/Mayur/Documents/SMU/Sem3/Visual Analytics/Session10/In-class_Ex_10/data/20news/"
```

# Read all files from folder into a dataframe

```{r}
read_folder <- function(infolder) {
  tibble(file = dir(infolder,
                    full.names = TRUE)) %>%
    mutate(text = map(file,
                      read_lines)) %>%
    transmute(id = basename(file),
              text) %>%
    unnest(text)
}
```

# Reading all the messages from the 20news folder

```{r}
raw_text <- tibble(folder=
                     dir(news20,
                         full.names = TRUE)) %>%
  mutate(folder_out = map(folder,
                          read_folder)) %>%
  unnest(cols = c(folder_out)) %>%
  transmute(newsgroup = basename(folder),
            id, text)
write_rds(raw_text, "C:/Users/Mayur/Documents/SMU/Sem3/Visual Analytics/Session10/In-class_Ex_10/data/rds/news20.rds")
```

# EDA to show the frequency of messages by newsgroup

```{r}
raw_text %>%
  group_by(newsgroup) %>%
  summarize(messages = n_distinct(id)) %>%
  ggplot(aes(messages, newsgroup)) +
  geom_col(fill = "lightblue") +
  labs(y = NULL)
```
# Removing header and automated email signatures

```{r}
cleaned_text <- raw_text %>%
  group_by(newsgroup, id) %>%
  filter(cumsum(text == "") > 0,
         cumsum(str_detect(
           text, "^--")) == 0) %>%
  ungroup()
```

# Removing lines with nested text representing quotes from other users

```{r}
cleaned_text <- cleaned_text %>%
  filter(str_detect(text, "^[^>]+[A-Za-z\\d]")
         | text == "",
         !str_detect(text,
                     "writes(:|\\.\\.\\.)$"),
         !str_detect(text,
                     "^In article <")
  )
```

# Text Data Processing (stop words removal and tokenize words)

```{r}
usenet_words <- cleaned_text %>%
  unnest_tokens(word, text) %>%
  filter(str_detect(word, "[a-z']$"),
         !word %in% stop_words$word)
```

# Count words by newsgroup

```{r}
words_by_newsgroup <- usenet_words %>%
  count(newsgroup, word, sort = TRUE) %>%
  ungroup()
```

# Visualizing words

```{r}
wordcloud(words_by_newsgroup$word,
          words_by_newsgroup$n,
          max.words = 300)
```
# Visualizing words in newsgroups


```{r}
set.seed(1234)

words_by_newsgroup %>%
  filter(n >0) %>%
ggplot(aes(label = word,
           size = n)) +
  geom_text_wordcloud() +
  theme_minimal() +
  facet_wrap(~newsgroup)
```

# Computing tf-idf within newsgroup

```{r}
tf_idf <- words_by_newsgroup %>%
  bind_tf_idf(word, newsgroup, n) %>%
  arrange(desc(tf_idf))
```

# Computing tf-idf within newsgroup

```{r}
DT::datatable(tf_idf, filter = 'top') %>%
  formatRound(columns = c('tf', 'idf',
                          'tf_idf'),
              digits = 3) %>%
  formatStyle(0,
              target = 'row',
              lineHeight='25%')
```

# Visualising tf-idf within newsgroups

```{r}
tf_idf %>%
  filter(str_detect(newsgroup, "^sci\\.")) %>%
  group_by(newsgroup) %>%
  slice_max(tf_idf,
            n = 12) %>%
  ungroup() %>%
  mutate(word = reorder(word,
                        tf-idf)) %>%
  ggplot(aes(tf_idf,
             word,
             fill = newsgroup)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~newsgroup,
             scales = 'free') +
  labs(x = 'tf-idf',
       y = NULL)
```

# Pairwise correlation between newsgroups

```{r}
newsgroup_cors <- words_by_newsgroup %>%
  pairwise_cor(newsgroup,
               word,
               n,
               sort = TRUE)
```


# visualising correlation as a network

```{r}
set.seed(2017)

newsgroup_cors %>%
  filter(correlation > .025) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation,
                     width = correlation)) +
  geom_node_point(size = 6,
                  color = "lightblue") +
  geom_node_text(aes(label = name),
                 color = "red",
                 repel = TRUE) +
  theme_void()
```
# Bigram

```{r}
bigrams <- cleaned_text %>%
  unnest_tokens(bigram,
                text,
                token = "ngrams",
                n = 2)
```

# Cleaning bigram

```{r}
bigrams_separated <- bigrams %>%
  filter(bigram != 'NA') %>%
  separate(bigram, c('word1','word2'),
  sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)
```

```{r}
bigram_counts <- bigrams_filtered %>%
count(word1, word2, sort = TRUE)
```

# Create a network graph from bigram data frame

```{r}
bigram_graph <- bigram_counts %>%
  filter(n > 3) %>%
  graph_from_data_frame()
bigram_graph
```

# Visualizing a network of bigrams with ggraph

```{r}
set.seed(1234)

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name),
                 vjust = 1,
                 hjust = 1)
```
# Improved visuvalization

```{r}
set.seed(1234)

a <- grid::arrow(type = "closed",
                 length = unit(0.15, "inches"))

ggraph(bigram_graph,
       layout = "fr") +
  geom_edge_link(aes(edge_alpha = n),
                 show.legend = FALSE,
                 arrow = a,
                 end_cap = circle(0.07, 'inches')) +
  geom_node_point(color = "lightblue",
                  size = 5) +
  geom_node_text(aes(label = name),
                 vjust = 1,
                 hjust = 1) +
  theme_void()
```


