---
title: "In-class Exercise 7"
description: |
  This post presents the step by step instructions that were followed to achieve the corresponding visualizations in Session 7 of our Visual Analytics Module.
author:
  - name: Mayurapriyann Arulmozhi Baskaran
    url: https://www.linkedin.com/in/mayurapriyann/
date: 06-26-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
    smooth_scroll: true
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE)
```


## Installing and Launching R Packages


```{r }
packages = c('DT','ggiraph','plotly','tidyverse')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

## Importing Data

```{r }
exam_data <- read_csv("C:/Users/Mayur//Documents/SMU/Sem3/Visual Analytics/Session7/In-class_Ex07/data/Exam_data.csv")
```


## Bar Plot

```{r }
ggplot(data = exam_data, aes(x = RACE)) +
  geom_bar()
```

## Dot Plot

```{r }
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_dotplot(binwidth = 2.5,dotsize = 0.5) +
  scale_y_continuous(NULL, breaks = NULL)
```


## Dot Plot (colored by RACE)

```{r }
ggplot(data = exam_data, aes(x = MATHS, fill = RACE)) +
  geom_dotplot(binwidth = 2.5,dotsize = 0.5) +
  scale_y_continuous(NULL, breaks = NULL)
```

# Histogram

```{r }
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, color = "black", fill = "light blue")
```

## Histogram (colored by GENDER)

```{r }
ggplot(data = exam_data, aes(x = MATHS, fill = GENDER)) +
  geom_histogram(bins = 20, color = "grey30")
```

# Box plot with jitter

```{r }
ggplot(data = exam_data,
       aes(y = MATHS,
           x = GENDER)) +
  geom_boxplot() +
  geom_point(position="jitter",
             size = 0.5)
```


# Interactive Plots (with tooltip)

```{r }
p <- ggplot(data = exam_data,
        aes(x = MATHS)) +
    geom_dotplot_interactive(
      aes(tooltip = ID),
      stackgroups = TRUE,
      binwidth = 1,
      method = "histodot") +
    scale_y_continuous(NULL, breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```

## Interactive Plots (hover effect with data_id aesthetic)

```{r }
p <- ggplot(data = exam_data,
        aes(x = MATHS)) +
    geom_dotplot_interactive(
      aes(data_id = CLASS),
      stackgroups = TRUE,
      binwidth = 1,
      method = "histodot") +
    scale_y_continuous(NULL, breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```


## Interactive Plots (hover effect with data_id aesthetic with tooltip)

```{r }
p <- ggplot(data = exam_data,
        aes(x = MATHS)) +
    geom_dotplot_interactive(
      aes(data_id = CLASS, 
          tooltip = CLASS),
      stackgroups = TRUE,
      binwidth = 1,
      method = "histodot") +
    scale_y_continuous(NULL, breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```
## Interactive Plots (onclick)

```{r }
exam_data$onclick <- sprintf("window.open(\"%s%s\")",
"https://www.moe.gov.sg/schoolfinder?journey=Primary%20school", as.character(exam_data$ID) )

p <- ggplot(data = exam_data,
        aes(x = MATHS)) +
    geom_dotplot_interactive(
      aes(onclick = onclick),
      stackgroups = TRUE,
      binwidth = 1,
      method = "histodot") +
    scale_y_continuous(NULL, breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
``` 

# Interactive visualization using plotly

```{r }
plot_ly(data = exam_data,
        x = ~ENGLISH,
        y = ~MATHS,
        color = ~RACE
        )

```

## Customising the colours

```{r }
pal <- c("red","blue","green","purple")

plot_ly(data = exam_data,
        x = ~ENGLISH,
        y = ~MATHS,
        color = ~RACE,
        colors = pal
        )

```

## Customising layout

```{r }
plot_ly(data = exam_data,
        x = ~ENGLISH,
        y = ~MATHS,
        text = ~paste("Student ID:", ID,
                      "<br>Class:", CLASS),
        color = ~RACE,
        colors = "Set1") %>%
layout(title = "English Score Versus Maths Score",
       xaxis = list(range = c(0,100)),
       yaxis = list(range = c(0,100)))

```

## Coordinated Multiple Views With Plotly

```{r }
d <- highlight_key(exam_data)
p1 <- ggplot(data = d,
             aes(x=MATHS,
                 y=ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
p2 <- ggplot(data = d,
             aes(x=MATHS,
                 y=SCIENCE)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
subplot(ggplotly(p1),
        ggplotly(p2))

```

## Interactive datatable

```{r }
DT::datatable(exam_data)
```

## Linked brushing: crosstalk method

```{r }
d <- highlight_key(exam_data)
p <- ggplot(d,
            aes(ENGLISH,
                MATHS)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
gg <- highlight(ggplotly(p),
                "plotly_selected")
crosstalk::bscols(gg,
                  DT::datatable(d),
                  widths=5)
```

