---
title: "In-class Exercise 8"
description: |
  This post presents the step by step instructions that were followed to achieve the corresponding visualizations in Session 8 of our Visual Analytics Module.
author:
  - name: Mayurapriyann Arulmozhi Baskaran
    url: https://www.linkedin.com/in/mayurapriyann/
date: 07-10-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
    smooth_scroll: true
    theme: united
    highlight: tango
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, 
                      echo = TRUE, 
                      eval = TRUE,
                      message = FALSE, 
                      error = FALSE, 
                      warning=FALSE)
```

```{r }
packages = c('igraph','tidygraph','ggraph','visNetwork','lubridate','clock',
             'tidyverse')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

```{r}
GAStech_nodes <- read_csv("C:/Users/Mayur/Documents/SMU/Sem3/Visual Analytics/Session9/In-class_Ex09/data/GAStech_email_node.csv")
GAStech_edges <- read_csv("C:/Users/Mayur/Documents/SMU/Sem3/Visual Analytics/Session9/In-class_Ex09/data/GAStech_email_edge-v2.csv")
```

```{r}
GAStech_edges$SentDate = dmy(GAStech_edges$SentDate)
GAStech_edges$Weekday = wday(GAStech_edges$SentDate,
                             label = TRUE,
                             abbr = FALSE)
```

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source, target, Weekday) %>%
    summarise(Weight = n()) %>%
  filter(source!=target) %>%
  filter(Weight > 1) %>%
  ungroup()
```

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated,
                           directed = TRUE)
```

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated,
                           directed = TRUE)
```

# Plotting a basic network graph

```{r}
ggraph(GAStech_graph) +
  geom_edge_link() +
  geom_node_point()
```
# Changing the default network graph theme

```{r}
g <- ggraph(GAStech_graph) +
  geom_edge_link(aes()) +
  geom_node_point(aes())
g + theme_graph()
```

# Fruchterman and Reingold layout

```{r}
g <- ggraph(GAStech_graph,
            layout = "fr") +
  geom_edge_link(aes()) +
  geom_node_point(aes())
g + theme_graph()
```


# Adding colours

```{r}
g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes()) +
  geom_node_point(aes(colour = Department,
                      size = 3))
g + theme_graph()
```

# Modifying edges

```{r}
g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes(width = Weight),
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
                      size = 3))
g + theme_graph()
```

# Working with facet_edges()

```{r}
set_graph_style()

g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes(width = Weight),
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
                      size = 2))
g + facet_edges(~Weekday)
```

# Working with facet_edges() change legend position

```{r}
set_graph_style()

g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes(width = Weight),
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
                      size = 2)) +
  theme(legend.position = 'bottom')
g + facet_edges(~Weekday)
```

# Framed facet graph

```{r}
set_graph_style()

g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes(width = Weight),
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
                      size = 2))
g + facet_edges(~Weekday) +
  th_foreground(foreground = "grey80",
                border = TRUE) +
  theme(legend.position = 'bottom')
  
```

# Working with facet_nodes()

```{r}
set_graph_style()

g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes(width = Weight),
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
                      size = 2))
g + facet_nodes(~Department) +
  th_foreground(foreground = "grey80",
                border = TRUE) +
  theme(legend.position = 'bottom')
  
```

# Computing centrality indices

```{r}
g <- GAStech_graph %>%
  mutate(betweenness_centrality = centrality_betweenness()) %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(width = Weight),
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
                      size = betweenness_centrality))
g + theme_graph()
```

# Computing centrality indices without mutate

```{r}
g <- GAStech_graph %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(width = Weight),
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
                      size = centrality_betweenness()))
g + theme_graph()
```

# Visualising Community

```{r}
g <- GAStech_graph %>%
  mutate(community = as.factor(group_edge_betweenness(weights = Weight, directed = TRUE))) %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(width = Weight),
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = community))
g + theme_graph()
```

# Data preparation
# (to plot interactive network graph)

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  left_join(GAStech_nodes, by = c("sourceLabel" = "label")) %>%
  rename(from = id) %>%
  left_join(GAStech_nodes, by = c("targetLabel" = "label")) %>%
  rename(to = id) %>%
  filter(MainSubject == "Work related") %>%
  group_by(from, to) %>%
    summarise(weight = n()) %>%
  filter(from!=to) %>%
  filter(weight > 1) %>%
  ungroup()
```

# Plotting interactive network graph

```{r}
visNetwork(GAStech_nodes,
            GAStech_edges_aggregated)
```

# Plotting interactive network graph without layout

```{r}
visNetwork(GAStech_nodes,
            GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr")
```

# Rename department to group

```{r}
GAStech_nodes <- GAStech_nodes %>%
  rename(group=Department)
```

# Visualising the nodes

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

# Visualising the nodes with Interactivity

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visOptions(highlightNearest = TRUE,
             nodesIdSelection = TRUE) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```